---
title: "Econometrics"
author: "Óskar, Andreas, Jonas"
date: "12/21/2020"
output: html_document
___
```{r}
library(tidyverse)
```

```{r}
#vehicles <- read.csv("rawdata/maindata.csv")

vdf <- vehicles[-c(1:5,16,21,22)]


#Here we cut off for above 1million and below 100 dollars.
vdf <- subset(vdf,vdf$price < 1000000)
vdf <- subset(vdf,vdf$price > 100)
#changing cylinders into factors
vdf$cylinders <- as.factor(factor(vdf$cylinders, levels = c( NA , "3 cylinders" , "4 cylinders", "5 cylinders", "6 cylinders", "8 cylinders", "10 cylinders", "12 cylinders" ), ordered = "FALSE"))
#changing condition to a factor.
vdf$condition <- as.factor(factor(vdf$condition, levels = c( NA ,"salvage", "fair", "good", "excellent", "like new", "new"), ordered = "TRUE"))
#Changing fuel to a factor
vdf$fuel <- as.factor(factor(vdf$fuel, levels = c( NA ,"gas", "diesel", "other", "hybrid", "electric"), ordered = "FALSE"))
levels(vdf$fuel)
#Changing title_status to factor
vdf$title_status <- as.factor(factor(vdf$title_status, levels = c( NA ,"clean", "lien", "missing", "salvage", "rebuilt", "parts only"), ordered = "FALSE"))
summary(vdf$cylinders)

navdf <- drop_na(vdf, (c("cylinders", "fuel", "condition", "title_status", "model")))



